### 2018-07-26 ai-assistants/paper.tex
What is an AI assistant?

\documentclass{article}
\usepackage{color}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{tgtermes}
\usepackage{forest}
\usepackage{fancyvrb}

\forestset{
  tt/.style={
    for tree={
      parent anchor=south,
      child anchor=north,
      grow=270,
      align=center,
      inner sep=0pt
    }
  },
  nice empty nodes up/.style={
    for tree={calign=fixed edge angles},
    delay={where content={}{shape=coordinate,for parent={for children={anchor=south}}}{}}
  }
}
\begin{document}
\title{What is an AI assistant?}
\author{Tomas Petricek}
\maketitle

\newcommand{\ident}[1]{\textnormal{\sffamily #1}}

\section{Introduction}
The idea of an \emph{AI assistant} has been discussed in many contexts in the AIDA project.
The Wrattler notebook system aims to provide a platform through which AI assistants can be 
easily used; the datadiff project developed an AI assistant which automates the process of 
reconciling inconsistencies between pairs of tabular datasets. However, we never clearly defined
what an AI assistant is. This document aims to collect some notes to help answer this question.
In the current version, these are mostly based on the needs of Wrattler and datadiff integration
in Wrattler -- as such, it clearly does not give the final answer, but merely one (limited, but
hopefully useful) perspective.

A simple way of trying to address the data wrangling problem would be to build a set of functions
that take some dirty datasets and produce one or more clean datasets. The user could then call these
functions on their dirty data and would get a clean result (conceivably, this could also be further 
automated). However, this simple approach has two problems. First, real-world data sets often 
contain numerous corner cases that require some human insight. Second, we want the data analysis
to be transparent and be able to understand how have the data wrangling problems been solved. 
A simple function does not address these because it does not encourage human interaction\footnote{At
best, users can run the function again with different parameters, but they get no suggestion about 
suitable values of the parameters.} and it is not transparent -- it produces a result without 
indicating how and why.

\section{What makes AI assistants unique}

AI assistants address the above issues as follows. First, AI assistants are interactive. Given 
some input data, they perform some analysis and provide the user with several options to choose 
from (those can be most likely solutions to some problem, possible ways of resolving ambiguous
corner cases, or a choice of next steps the AI assistant can follow). The options can be a finite
list (where the user has to choose one) or open-ended (where the user can specify some input).
They should also be sorted by likelihood (if we always choose the most likely option, the AI
assistant will behave as a fully automatic function).

Second, AI assistants do not directly transform data, but instead, generate scripts (or, more 
formally, expressions) in simple languages (domain-specific languages, or DSLs) that define how 
data should be transformed. The AI assistant can still use sophisticated machine learning 
algorithms (that do not produce an explanation of how they work) behind the scenes, but the end
result will always be a simple script that can be reviewed and understood by the user.

Note that each AI assistant can use its own domain-specific language -- the language should be
as small as possible, i.e. it should be able to express the data transformations that the AI
assistant can recommend, but nothing more.

The combination of interactivity and domain-specific languages allows us to do a number of 
interesting things:

\begin{itemize}
\item Scripts generated by AI assistant can be translated to R or Python (and it is easy to 
  add support for other languages), which means that one can use an AI assistant during data
  analysis and exploration, but then end up with code that can be used in production in any
  programming language.

\item The interactive nature of AI assistants makes it possible to give immediate feedback to
  users and let them correct potential issues early. For example, in Wrattler, we display the
  options that AI assistant provides together with a preview of results (of applying the
  generated expression to a sample dataset). 
\end{itemize}

\section{AI assistants as trees}
We can think of an AI assistant as a function that takes some input data and generates
a tree where each node offers a finite list of options that the user can choose from. (For
simplicity, we ignore the possibility of open-ended options where the user specifies some input 
and we also ignore the fact that the AI assistant might take some initial parameters in addition
to the input). Formally:
%
\begin{equation*}
\begin{array}{rcll}
a &=& \ident{Data} \rightarrow t &\textnormal{(AI assistant)}\\
t &=& (n_1, p_1, t_1), \ldots (n_k, p_k, e_k, t_k) &\textnormal{(Provided options)}\\
\mathit{eval} &=& \ident{Data}\times e \rightarrow \ident{Data}&\textnormal{(Evaluator for the DSL)}
\end{array}
\end{equation*}
%
An AI assistant $a$ is a function that takes some input data and returns a root tree node $t$. 
A tree node contains a list of options $(n_i, p_i, e_i, t_i)$ where $n_i$ is a name of the node (shown
to the user, but otherwise unimportant), $p_i$ is a number that gives the likelihood of the 
option (between 0 and 1), $e_i$ is an expression in the domain-specific language that represents
the data transformation attached to the node and $t_i$ is a sub-tree that can be empty or can offer 
more follow-up options. Note that this sub-tree can be generated lazily -- it is only needed if the 
user chooses the option with which it is associated and so we do not need to evaluate the full tree 
upfront.

The nature of expressions $e$ depends on the particular AI assistant and so we leave that abstract
for now. However, we certainly need to be able to apply the transformation on sample dataset, 
which is captured by the $\mathit{eval}$ function, which takes data together with a DSL expression
and returns transformed data. An (abstract) example tree might look as follows:

\vspace{0.5em}
\begin{forest}, tt  
  [
    [{$\mathbf{n_1}$\\$p=0.9$\\$e=e_1$}
      [{$\quad\mathbf{n_{11}}\quad$\\$p=0.9$\\$e=e_{11}$}]
      [{$\quad\mathbf{n_{12}}\quad$\\$p=0.7$\\$e=e_{12}$}]
    ]
    [{$\mathbf{n_2}$\\$p=0.8$\\$e=e_2$}
      [{$\quad\mathbf{n_{21}}\quad$\\$p=0.9$\\$e=e_{21}$}]
      [{$\quad\mathbf{n_{22}}\quad$\\$p=0.8$\\$e=e_{22}$}]
    ]
    [{$\mathbf{n_3}$\\$p=0.7$\\$e=e_3$}
      [{$\quad\mathbf{n_{31}}\quad$\\$p=0.8$\\$e=e_{31}$}]
      [{$\quad\mathbf{n_{32}}\quad$\\$p=0.6$\\$e=e_{32}$}]
    ]
  ]
\end{forest}
\vspace{0.5em}

\noindent
Here, the root offers three options represented by nodes $n_1, n_2, n_3$. The most likely one
as recommended by the AI assistant is $n_1$. The expression attached to the node is $e_1$ and
when the user selects the node, we can display a preview using sample data $\mathit{sample}$
by evaluating $\mathit{eval}(\mathit{sample}, e_1)$.

Assuming the user chooses the most likely option $n_1$, they will then be presented with two
subsequent options, $n_{11}, n_{12}$. Even though the AI assistant marks the first node as more
likely, the user may review the previews $\mathit{eval}(\mathit{sample}, e_{11})$,
$\mathit{eval}(\mathit{sample}, e_{12})$ and decide that the latter is actually the correct
data cleaning operation. They can then choose the node $n_{12}$ as the final one use the 
cleaning script represented by the DSL expression $e_{12}$.

\section{Examples of AI assistants}

What would a sample AI assistant look like when seen through the above model? We consider two
examples. The first is hypothetical AI assistant for extracting data from CSV files. The second
is the datadiff assistant. For datadiff, we discuss the current implementation in Wrattler, but
also briefly sketch a possible improvement (that improves the interactivity aspect).

\subsection{Cleaning CSV files}

As an example, consider the following broken CSV file. The fle starts with two meta-data lines
(one specifying the name and another specifying copyright information). This is followed by headers
and actual data.

\begin{Verbatim}[numbers=left,xleftmargin=8mm]
Counts of things
Copyright (c) Poor data export company, London, UK
Name, City, Count
Joe, London, 3
Jane, Edinburgh, 16
Jim, Cardiff, na
\end{Verbatim}

A hypothetical minimalistic AI assistant for cleaning CSV files could generate tree with two
levels of nodes. In the first level, it will aim to infer where the actual data is in the CSV
file (and it will offer its best guesses to the user). In the second level, it will aim to infer
the types of the columns.

The domain-specific language of data cleaning operations consists of two functions named
$\mathit{range}$ and $\mathit{types}$ that specify the range of data in the file and the types
of the columns: $\mathit{range}(\mathit{lines}=l_n\ldots l_m, \mathit{cols}=k)$ denotes that
the data starts on line $l_n$ and ends on line $l_m$ and there are $k$ columns; 
$\mathit{types}(t_1,\ldots, t_k)$ specifies that the types of the $k$ columns are 
$t_1, \ldots, t_k$. The types are either $\ident{string}$ or $\ident{int}$.

Given the above file as an input, the AI assistant might produce the following tree:
%
\begin{equation*}
\begin{array}{l}
p=0.9,~ e=\mathit{range}(\mathit{lines}=2\ldots 6, \mathit{cols}=3)\\[0.2em]
\qquad\longrightarrow p=0.9,~ e=\mathit{range}(\mathit{lines}=2\ldots 6, \mathit{cols}=3); \mathit{types}(\ident{string},\ident{string},\ident{string})\\[0.2em]
\qquad\longrightarrow p=0.3,~ e=\mathit{range}(\mathit{lines}=2\ldots 6, \mathit{cols}=3); \mathit{types}(\ident{string},\ident{string},\ident{int})\\[0.4em]
p=0.7,~ e=\mathit{range}(\mathit{lines}=3\ldots 6, \mathit{cols}=3)\\[0.2em]
\qquad\longrightarrow p=0.8,~ e=\mathit{range}(\mathit{lines}=3\ldots 6, \mathit{cols}=3); \mathit{types}(\ident{string},\ident{string},\ident{int})\\[0.2em]
\qquad\longrightarrow p=0.7,~ e=\mathit{range}(\mathit{lines}=3\ldots 6, \mathit{cols}=3); \mathit{types}(\ident{string},\ident{string},\ident{string})\\[0.4em]
p=0.3,~ e=\mathit{range}(\mathit{lines}=1\ldots 6, \mathit{cols}=1)\\[0.2em]
\qquad\longrightarrow p=0.9,~ e=\mathit{range}(\mathit{lines}=1\ldots 6, \mathit{cols}=1); \mathit{types}(\ident{string})\\[0.2em]
\end{array}
\end{equation*}
%
At the first level, the AI assistant gives us three options. The first, most likely, one skips just 
the first line and identifies that there are three columns (perhaps because the second line contains 
two commas and can thus be interpreted as data row). The second skips first two lines and recognises
three columns. The AI assistant might find this less likely, but it is the option we want to choose
as users. Finally, the third option takes all lines, but then it has to join all columns into one
which is unlikely.

The user can choose the second option, at which point a user interface can evaluate the 
expression $\mathit{range}(\mathit{lines}=3\ldots 6, \mathit{cols}=3)$ and show a preview of 
the cleaning result. Next, the user has a choice of two ways of inferring types -- the most 
likely one (and the right one) is that the third column has a type \ident{int} (which means that
the value {\texttt na} in the last row will be treated as missing).

The resulting expression 
$\mathit{range}(\mathit{lines}=3\ldots 6, \mathit{cols}=3); \mathit{types}(\ident{string},\ident{string},\ident{int})$
can then be applied to the input data using the $\mathit{eval}$ function, which gives us 
columns \ident{Name} of type \ident{string}, \ident{City} of type \ident{string}
and \ident{Count} of type \ident{int} with data:
%
\begin{equation*}
\begin{array}{l}
(\textnormal{\sffamily "\hspace{-0.2em}Joe"}, \textnormal{\sffamily "\hspace{-0.2em}London"}, 3)\\
(\textnormal{\sffamily "\hspace{-0.2em}Jane"}, \textnormal{\sffamily "\hspace{-0.2em}Edinburgh"}, 16)\\
(\textnormal{\sffamily "\hspace{-0.2em}Jim"}, \textnormal{\sffamily "\hspace{-0.2em}Cardiff\hspace{0.1em}"}, \mathit{NaN})
\end{array}
\end{equation*}

\subsection{Datadiff (now)}

\subsection{Datadiff (better)}

\section{Evaluating AI assistants}

\subsection{Overall}
Following the best probability, we get the best result 
(this is equivalent to running a function)

\subsection{Interactivity}

Imagine there are some annomalies that we cannot detect.
Does the AI assistant provide an easy way for the user to address them?

\end{document}